cardinalR: Generating interesting high-dimensional data structures

### 2025-05-05 paper-cardinalR.Rmd
Added sphere section

#### Geometric Structure

- A **large uniform sphere** centered at the origin:
  - \(n_1\) points are sampled **uniformly on the surface** of a \(p\)-dimensional sphere of radius \(r_1\).
  - Generated using `gen_unifsphere(n_1, p, r_1)`.

- \(k\) **smaller uniform spheres**, each with:
  - \(n_2\) points sampled uniformly on a sphere of radius \(r_2\),
  - Positioned at different random locations in \(p\)-space:
    - Each center is drawn from \( \mathcal{N}(0, \texttt{loc}^2 I_p) \),
    - Ensures separation between clusters.

All points on all spheres are generated using the standard hyperspherical method:

- For each point:
  - \(u \sim U(-1, 1)\), the cosine of the polar angle,
  - \(\theta \sim U(0, 2\pi)\), the azimuthal angle (for 3D),
  - Generalized to \(p\)-dimensions to uniformly sample on the sphere’s surface.

>>>>

A large uniform sphere centered at the origin is created by sampling $n_1$ points uniformly on the surface of a $p$-dimensional sphere with a radius of $r_1$. The sampling is executed using the function `gen_unifsphere(n_1, p, r_1)`, which generates the desired points in the specified dimensional space.

In generation of $k$ smaller uniform spheres, each sphere contains $n_2$ points that are sampled uniformly on a sphere with a radius of $r_2$. These spheres are positioned at distinct random locations in $p$-space, with the center of each sphere being drawn from a normal distribution $\mathcal{N}(0, \texttt{loc}^2 I_p)$.

Points on spheres are generated using the standard hyperspherical method, which involves sampling $u \sim U(-1, 1)$ to determine the cosine of the polar angle, and sampling $\theta \sim U(0, 2\pi)$ to determine the azimuthal angle (for 3D).
